This paper aims to develop a statistical model to estimate the testing time needed before a piece of software is ready for release. MathWorks is a software company that develops Matlab and Simulink used for numerical calculations and simulations by many scientists and engineers. Having developed a new statistics package named StatWorks, the manager at MathWorks are interested in the time needed for testing before the software is ready for release. Since the number of bugs in StatWorks is unknown, it is impossible to determine whether the software is truly free of bugs. Having said that, MathWorks needs to release StatWorks knowing that it is sufficiently reliable. As an example, this paper requires that the probability of the software running without failure for 4 CPU hours is 95\%, however its parameters can be easily changed to fit other requirements.

In Section \ref{sec:simple-model} a simplified version of the model is developed, using the assumption that there are only 2 bugs in the software to develop the intuition for the model. In Section \ref{sec:2.2} an unknown number of bugs $u$ is introduced and some useful equations are derived. Section \ref{sec:estimators} uses the failure times from Mathworks in Appendix \ref{table-failure} and the method of Maximum Likelihood Estimators (MLE) to find estimates for the parameters of the model. Lastly, Section \ref{sec:results} introduces some assumptions about the development teams in MathWorks to estimate the time before StatWorks is sufficiently reliable and ready for release.